Multidimensional analysis and detection of informative features in human white matter

### 2021-01-25 reports/manuscript/paper.tex
Add nested cross-val figure. Update figures and text to remove underscores in bundle names. Add text to describe bundle profiles in supplementary material.
Mostly work in results section.

%% ****** Start of file apstemplate.tex ****** %
%%
%%   This file is part of the APS files in the REVTeX 4 distribution.
%%   Version 4.1r of REVTeX, August 2010
%%
%%
%%   Copyright (c) 2001, 2009, 2010 The American Physical Society.
%%
%%   See the REVTeX 4 README file for restrictions and more information.
%%
%
% This is a template for producing manuscripts for use with REVTEX 4.0
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.
%
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
\documentclass[10pt,%
               aps,%
               prl,%
               preprint,%
               superscriptaddress,%
               preprintnumbers,%
               amsmath,%
               floatfix,%
               endfloats*]{revtex4-2}
\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc} 
\usepackage[export]{adjustbox}  % Used to adjust figure frames
\usepackage{subcaption}
\usepackage{url}
\usepackage{color}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dcolumn}
\usepackage{wrapfig}
\usepackage{physics}
\usepackage[inline]{enumitem}
\usepackage[per-mode=symbol,group-separator={,},group-minimum-digits=4]{siunitx}
\usepackage{todonotes}
\usepackage{graphicx}
\graphicspath{ {../paper_figures} }

@@

\begin{acknowledgments}
    {\it Acknowledgments.-}
    This work was supported by BRAIN Iniative grant 1RF1MH121868-01 from the
    National Institutes for Mental Health and by a grant from the Gordon \& Betty
    Moore Foundation and the Alfred P. Sloan Foundation to the University of
    Washington eScience Institute Data Science Environment. We would like to thank
    Scott Murray for a useful discussion of the SGL method and Mareike Grotheer for
    helpful comments on the manuscript.
\end{acknowledgments}

>>>>

%% ****** Start of file apstemplate.tex ****** %
%%
%%   This file is part of the APS files in the REVTeX 4 distribution.
%%   Version 4.1r of REVTeX, August 2010
%%
%%
%%   Copyright (c) 2001, 2009, 2010 The American Physical Society.
%%
%%   See the REVTeX 4 README file for restrictions and more information.
%%
%
% This is a template for producing manuscripts for use with REVTEX 4.0
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.
%
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
\documentclass[10pt,%
               aps,%
               prl,%
               preprint,%
               superscriptaddress,%
               preprintnumbers,%
               amsmath,%
               floatfix,%
               endfloats*]{revtex4-1}
\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage[export]{adjustbox}  % Used to adjust figure frames
\usepackage{subcaption}
\usepackage{url}
\usepackage{color}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dcolumn}
\usepackage{wrapfig}
\usepackage{physics}
\usepackage[inline]{enumitem}
\usepackage[per-mode=symbol,group-separator={,},group-minimum-digits=4]{siunitx}
\usepackage{todonotes}
\usepackage{graphicx}
\graphicspath{ {../paper_figures} }

@@

\begin{acknowledgments}
    {\it Acknowledgments.-}
    This work was supported by BRAIN Iniative grant 1RF1MH121868-01 from the
    National Institutes for Mental Health, by a grant from the Gordon \& Betty
    Moore Foundation and the Alfred P. Sloan Foundation to the University of
    Washington eScience Institute Data Science Environment, and by the Google Cloud Platform Academic Research Credits Program.
    We would like to thank
    Scott Murray for a useful discussion of the SGL method and Mareike Grotheer for
    helpful comments on the manuscript.
\end{acknowledgments}

### 2021-01-25 reports/manuscript/sections/conclusion.tex
Mostly work in results section.

We present here a novel method for analysis of dMRI tractometry data that
relies on the Sparse Group Lasso \cite{simon2013sparse} to (a) make accurate
predictions of phenotypic properties of individual subjects while,
simultaneously, (b) identifying the features of the white matter that are
most important for this prediction in a completely data-driven approach. The
method is broadly applicable to a wide range of research questions: it
performs well in predicting both continuous variables, such as biological
age, as well as categorical variables, such as whether a person is a patient
or a healthy control. In both of these cases, SGL out-performs previous
algorithms that have been developed for these tasks
\cite{sarica2017corticospinal, Richard2018-ux, mcpherson2020single}. The
nested cross-validation approach used to fit the model and make both
predictions and inferences from the model guards against overfitting and
tunes the degree of sparseness required by the algorithm. This means that SGL
can accurately describe phenomena that are locally confined to a particular
anatomical location or diffusion property (e.g., FA in the CST) as well as
phenomena that are widely distributed amongst brain regions and measured
diffusion properties.

Specifically, we demonstrated that the algorithm correctly detects the fact
that ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a targeted ROI-based approach
\cite{sarica2017corticospinal}. In contrast, for the analysis of biological
age, the coefficients identified by the algorithm are very widely distributed
across many parts of the white matter, mirroring previous results that show a
large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.

@@

Taken together, these results demonstrate the promise of the
group-regularized regression approach. Even at the scale of dozens of
subjects, the results provided by SGL are both accurate, as well as
interpretable \cite{Murdoch2019-ax}: tractometry capitalizes on domain
knowledge to engineer meaningful features; SGL scores these features based on
their relative importance; enables a visualization of these feature
importance scores in the anatomical coordinate frame of the bundles (e.g.,
\cref{fig:class-results:coefs3d,fig:age-results:wh-coefs3d,fig:age-results:hbn-coefs3d,fig:age-results:cc-coefs3d})
and provides a means to understand model errors (e.g.,
\cref{fig:class-results:false-positives,fig:class-results:false-negatives}) .
Thus, this multivariate analysis approach both (a) achieves high
cross-validated accuracy for precision medicine applications of dMRI data and
(b) identifies relevant features of brain anatomy that can further our
neuroscientific understanding of clinical disorders.

Neuroscience has entered an era in which consortium efforts are amassing
large datasets of high-quality dMRI measurements to address a variety of
scientific questions \cite{jernigan2016ping, jernigan2018abcd,
alexander2017open, Miller2016-hw, VanEssen2012}. The volume and complexity of
these data pose a substantial challenge. Dimensionality reduction with
tractometry, followed by analysis with the approach we present here promises to
capitalize on the wealth of data and on the co-measurement of interesting and
important phenotypical data about brain health and about the participants'
cognitive abilities. We also expect the group-regularized approach to improve
with larger datasets.

SGL has many other potential applications in neuroscience, because
of the hierarchical and grouped nature of many data types that are
collected in multiple sample points within anatomically-defined areas.
For example, this method may be useful to understand the relationship
between fMRI recordings and behavior, where activity in each voxel
may co-vary with voxels within the same anatomical region and form
features and groups of features. Similarly, large-scale multi-electrode
recordings of neural activity in awake behaving animals are becoming
increasingly feasible \cite{steinmetz2018distributed, Jun2017-gv} and
these recordings can form features (neurons) and groups (neurons within
an anatomical region). More ambitiously perhaps, this approach may be
used to understand the role of correlations in so-called resting-state
fMRI time-series and behavior, where pairwise correlations between
voxels in different anatomical regions are features in the matrix and
features may be grouped by pairs of anatomical regions. Given the large
number of voxels in the surface of the gray matter and given that
correlations increase the number of features by a factor of $n^2$, this
would pose a challenging problem to solve using SGL.

@@

The method is packaged as open-source software called AFQ-Insight
that is openly available, and provides a clear API to allow
for extensions of the method. The sofware integrates within a
broader automated fiber quantification software ecosystem: AFQ
\cite{yeatman2012tract}, which extracts tractometry data from raw and
processed dMRI datasets, as well as AFQ-Browser, which visualizes
tractometry data and facilitates sharing of the results of dMRI studies
\cite{yeatman2018browser}. To facilitate reproducibility and ease use
of the software, the results presented
in this paper are also provided in
\url{https://github.com/richford/AFQ-Insight/tree/master/examples/preprint-notebooks} as a series of Jupyter notebooks
\cite{kluyver2016jupyter}.

>>>>

We present here a novel method for analysis of dMRI tractometry data that
relies on the Sparse Group Lasso \cite{simon2013sparse} to make accurate
predictions of phenotypic properties of individual subjects while,
simultaneously, identifying the features of the white matter that are
most important for this prediction. The method is data-driven and it is broadly applicable to a wide range of research questions: it
performs well in predicting both continuous variables, such as biological
age, as well as categorical variables, such as whether a person is a patient
or a healthy control. In both of these cases, SGL out-performs previous
algorithms that have been developed for these tasks
\cite{sarica2017corticospinal, Richard2018-ux, mcpherson2020single}. The
nested cross-validation approach used to fit the model and make both
predictions and inferences from the model guards against overfitting and
tunes the degree of sparseness required by the algorithm. This means that SGL
can accurately describe phenomena that are locally confined to a particular
anatomical location or diffusion property (e.g., FA in the CST) as well as
phenomena that are widely distributed amongst brain regions and measured
diffusion properties.

Specifically, we demonstrated that the algorithm correctly detects the fact
that ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a bundle-of-interest approach
\cite{sarica2017corticospinal}. In contrast, for the analysis of biological
age, the coefficients identified by the algorithm are very widely distributed
across many parts of the white matter, mirroring previous results that show a
large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.

@@

Taken together, these results demonstrate the promise of the
group-regularized regression approach. Even at the scale of dozens of
subjects, the results provided by SGL are both accurate, as well as
interpretable \cite{Murdoch2019-ax}: tractometry capitalizes on domain
knowledge to engineer meaningful features; SGL scores these features based on
their relative importance; enables a visualization of these feature
importance scores in the anatomical coordinate frame of the bundles (e.g.,
\cref{fig:class-results:coefs3d,fig:age-results:wh-coefs3d,fig:age-results:hbn-coefs3d,fig:age-results:cc-coefs3d})
and provides a means to understand model errors (e.g.,
\cref{fig:class-results:false-positives,fig:class-results:false-negatives}) .
Thus, this multivariate analysis approach achieves high
cross-validated accuracy for precision medicine applications of dMRI data and identifies relevant features of brain anatomy that can further our
neuroscientific understanding of clinical disorders.

Neuroscience has entered an era in which consortium efforts are amassing
large datasets of high-quality dMRI measurements to address a variety of
scientific questions \cite{jernigan2016ping, jernigan2018abcd,
alexander2017open, Miller2016-hw, VanEssen2012}, but the volume and
complexity of these data pose substantial challenges. Tractometry
followed by analysis with SGL provide a promising approach to
distill meaningful insights from the wealth of data measured in these efforts.
%We also expect the group-regularized approach to improve
% with larger datasets. <= ARIEL : I feel like we've demonstrated that already
% with the age regression results

SGL has many other potential applications in neuroscience, because
of the hierarchical and grouped nature of many data types that are
collected in multiple sample points within anatomically-defined areas.
For example, this method may be useful to understand the relationship
between fMRI recordings and behavior, where activity in each voxel
may co-vary with voxels within the same anatomical region and form
features and groups of features. Similarly, large-scale multi-electrode
recordings of neural activity in awake behaving animals are becoming
increasingly feasible \cite{steinmetz2018distributed, Jun2017-gv} and
these recordings can form features (neurons) and groups (neurons within
an anatomical region).

% More ambitiously perhaps, this approach may be
% used to understand the role of correlations in so-called resting-state
% fMRI time-series and behavior, where pairwise correlations between
% voxels in different anatomical regions are features in the matrix and
% features may be grouped by pairs of anatomical regions. Given the large
% number of voxels in the surface of the gray matter and given that
% correlations increase the number of features by a factor of $n^2$, this
% would pose a challenging problem to solve using SGL.

@@

The method is packaged as open-source software called AFQ-Insight
that is openly available (ARIEL: ADD URL), and provides a clear API to allow
for extensions of the method. The sofware integrates within a
broader automated fiber quantification software ecosystem: AFQ
\cite{yeatman2012tract} and pyAFQ (https://yeatmanlab.github.io), which extract tractometry data from raw and processed dMRI datasets, as well as AFQ-Browser, which visualizes tractometry data and facilitates sharing of the results of dMRI studies \cite{yeatman2018browser}. To facilitate reproducibility and ease use of the software, the results presented
in this paper are also provided in \url{https://github.com/richford/AFQ-Insight/tree/master/examples/preprint-notebooks} as a series of Jupyter notebooks
\cite{kluyver2016jupyter}.

### 2021-01-25 reports/manuscript/sections/results.tex
Mostly work in results section.

\begin{figure*}
    \vspace{3cm}
    \rlap{
        \includegraphics[width=0.33\textwidth]{als_coefs.pdf}
    }
    \vspace{-8.4cm}
    \includegraphics[width=\textwidth]{als_results.pdf}
    {\phantomsubcaption\label{fig:class-results:probs}}
    {\phantomsubcaption\label{fig:class-results:coefs3d}}
    {\phantomsubcaption\label{fig:class-results:tract-profiles}}
    {\phantomsubcaption\label{fig:class-results:false-positives}}
    {\phantomsubcaption\label{fig:class-results:false-negatives}}
    \caption{%
        {\bf SGL accurately and interpretably predicts ALS diagnosis.}
        \label{fig:class-results}
        {\bf (a)} Classification probabilities for ALS diagnosis, with
        controls on the left, patients on the right, predicted controls in
        blue, and predicted patients in orange. So orange dots on the left
        represent false positives, while blue dots on the right represent
        false negatives. We achieve {\protect\alsAccuracy}\% accuracy with an
        ROC AUC of {\protect\alsRocAuc}.
        {\bf (b)} SGL coefficients are presented on the core fibers of major
        fiber bundles. They exhibit high group sparsity and are concentrated
        in the FA of the CST. The brain is oriented with the right hemisphere
        in the foreground and anterior to the right of the page. The CSTL,
        CSTR, and CFA bundles are indicated for orientation.
        {\bf (c)} SGL identifies three portions of the CST as important,
        where $\hat{\beta}$ (dashed line, right axis) has large values. These
        are centered around nodes 30, 65, and 90, corresponding to locations
        of substantial differences in FA between the ALS and control groups
        (shaded areas indicates standard error of the mean).
        {\bf (d)} Bundle profiles for false positive classifications. Line
        colors correspond to the marker edge color in the top left plot.
        These individuals have reduced FA in the CST portions which SGL
        identified as important. Their misclassification is coherent with the
        feature importance and the group differences in FA.
        {\bf (e)} Individual bundle profiles for false negative
        classifications. These individuals have bundle profiles which
        oscillate between the group means.
    }
\end{figure*}

We developed a method for analyzing dMRI tractometry data with SGL. We
demonstrate the use of this method on four previously published datasets in
both a classification setting and a regression setting.

@@

Using data from a previous study of the corticospinal tract (CST) profile and
ALS \cite{sarica2017corticospinal}, we tested the performance of SGL in a
classification setting. The previous study predicted ALS status with a mean
accuracy of 80\% using a random forest algorithm based on a priori selection
of features within the CST. SGL delivers competitive
predictive performance, with a cross-validated accuracy of {\alsAccuracy}\%
and an area under the receiver operating characteristic curve (ROC AUC) of
{\alsRocAuc}, all without the need for a priori feature engineering. The
results of the classification prediction are shown in
\cref{fig:class-results:probs} with ``ground-truth'' ALS status separated
into columns, and predicted ALS status encoded by color. In addition to this
classification performance, SGL also identifies the white matter tracts most
important for ALS classification. The relative importance of white matter
features is captured in the $\beta$ coefficients from \cref{supp:eq:sgl} of
the supplemental material \cite{supplement}.
\Cref{fig:class-results:tract-profiles} depicts these coefficients along the
right CST, plotted over the FA values for the control and ALS subject groups.
We find that SGL selects FA metrics in the corticospinal tract and
particularly in the right corticospinal tract as most important to ALS
classification, confirming previous findings \cite{van2011upper,
toosy2003diffusion, sarica2014tractography, sage2007quantitative,
sage2009quantitative, karlsborg2004corticospinal, ellis1999diffusion,
cosottini2005diffusion, ciccarelli2009investigation, abe2010voxel} and
identifying the portions of the brain that were selected \emph{a priori} in
the previous study from which we collected our data
\cite{sarica2017corticospinal}.

The $\beta$ coefficients exhibit high bundle level sparsity; only some
bundles are important, which can be be confirmed by observing the value of
$\alpha$, the regularization hyperparameter that controls the mixture of the
group lasso and lasso penalties
(see \cref{supp:eq:sgl} of the supplemental material \cite{supplement}).
If $\alpha$ is closer to zero, it indicates that the phenotype in question
preferentially correlates with only a few groups of covariates. For the ALS
dataset, $\alpha = $ \alsLRatio, confirming that the white matter correlates
of ALS reside mostly in one bundle, namely the CST.

Analyzing the ways in which the model mislabels individuals also provides
insight. We found that mislabelled subjects are outliers relative to their
group with respect to diffusion features of the CST.
\Cref{fig:class-results:false-positives,fig:class-results:false-negatives}
depicts the group FA values along with FA values of mislabelled subjects,
four false negatives and three false-positives. The false positive
classifications have reduced FA in one or more of the three sections of the
CST where $\norm*{\hat{\beta}}$ is large
\cref{fig:class-results:tract-profiles}. The false negative subjects have FA
profiles that oscillate between the two group means. Thus, the SGL method
fails comprehensibly.

@@

To test the performance of SGL with tractometry data in a continuous
regression task, we focus here on the prediction of biological age in three
datasets named WH, HBN, and Cam-CAN (see supplemental material
\cite{supplement} for a description of each dataset). Prediction of ``brain
age'' is a commonly undertaken task in neuroimaging machine learning, in part
because these predictions, and deviations therefrom, may be diagnostic of
overall brain health (for a review, see Cole et al.~\cite{Cole2019-rz}).
However, as Nelson et al.~\cite{nelson2019biomarkers} have observed, aging
biomarkers are subject to unique challenges and tend to be noncausative. Our
interest in aging here relies on its utility as a methodological benthmark.
Biological age operates on a natural scale, with meaningful and easily
understood units, and it's popularity as a machine learning target makes it
valuable for comparisons with other studies.

The WH \cite{yeatman2014lifespan}, HBN \cite{alexander2017open}, and Cam-CAN
\cite{shafto2014cambridge,taylor2017cambridge} datasets used here contain
data from 76, 978, and 640 subjects, respectively, ranging from 6-50, 5-21,
and 18-88 years of age, respectively. In each case, biological age was used
as the predicted variable $y$. SGL was fit to the tractometry-extracted
features FA and MD in 18 major brain tracts, with diffusion metrics extracted
from diffusion tensor imaging (DTI) for the WH dataset and diffusion kurtosis
imaging (DKI) \cite{jensen2005diffusion} the HBN and Cam-CAN datasets.

To evaluate the fit of the model, we used a nested cross-validation
procedure. In this procedure, batches of subjects are held out. For each
batch (or fold), the model is fully fit without this data. Then, once the
parameters are fixed, the model is inverted to predict the ages of held out
subjects based on the linear coeffiecients. This scheme automatically finds
the right level of sparsity and fits the coefficients to the ill-posed linear
model, while guarding against overfitting. SGL accurately predicts the age of
the subjects in this procedure, with a median absolute error of {\whMae},
{\hbnMae}, and {\camcanMae} years for the WH, HBN, and Cam-CAN datasets,
respectively and coefficients of determination $R^2 = $ {\whRsq} , {\hbnRsq},
and {\camcanRsq}, respectively (see \cref{fig:age-results}, top panels). The
predictions for Cam-CAN are competitive with a recent state of the art
prediction \cite{mcpherson2020single}, which used streamline density to
estimate the brain's structural connectivity and achieved $R^2 = 0.63$. The
median absolute errors are also lower than the results of a recent study that
predicted age in a large sample, based on diffusion MRI features
\cite{Richard2018-ux}.

>>>>

\begin{figure*}
    \vspace{3cm}
    \rlap{
        \includegraphics[width=0.33\textwidth]{als_coefs.pdf}
    }
    \vspace{-8.4cm}
    \includegraphics[width=\textwidth]{als_results.pdf}
    {\phantomsubcaption\label{fig:class-results:probs}}
    {\phantomsubcaption\label{fig:class-results:coefs3d}}
    {\phantomsubcaption\label{fig:class-results:tract-profiles}}
    {\phantomsubcaption\label{fig:class-results:false-positives}}
    {\phantomsubcaption\label{fig:class-results:false-negatives}}
    \caption{%
        {\bf SGL accurately and interpretably predicts ALS diagnosis.}
        \label{fig:class-results}
        {\bf (a)} Classification probabilities for ALS diagnosis, with
        controls on the left, patients on the right, predicted controls in
        blue, and predicted patients in orange. This is, orange dots on the left
        represent false positives, while blue dots on the right represent
        false negatives. We achieve {\protect\alsAccuracy}\% accuracy with an
        ROC AUC of {\protect\alsRocAuc}.
        {\bf (b)} SGL coefficients are presented on the core fibers of major
        fiber bundles. They exhibit high group sparsity and are concentrated
        in the FA of the CST. The brain is oriented with the right hemisphere
        in the foreground and anterior to the right of the page. The CSTL,
        CSTR, and CFA bundles are indicated for orientation.
        {\bf (c)} SGL identifies three portions of the CST as important,
        where $\hat{\beta}$ (dashed line, right axis) has large values. These
        are centered around nodes 30, 65, and 90, corresponding to locations
        of substantial differences in FA between the ALS and control groups
        (shaded areas indicates standard error of the mean).
        {\bf (d)} Bundle profiles for false positive classifications. Line
        colors correspond to the marker edge color in the top left plot.
        These individuals have reduced FA in the CST portions which SGL
        identified as important. Their misclassification is coherent with the
        feature importance and the group differences in FA.
        {\bf (e)} Individual bundle profiles for false negative
        classifications. These individuals have bundle profiles which
        oscillate between the group means.
    }
\end{figure*}

We developed a method for analyzing dMRI tractometry data that uses the Sparse
Group Lasso (SGL) to select features that are sparse both at the group (bundle) level, as well as overall.
We demonstrate the use of this method on four different datasets in
both a classification setting and a regression setting.

@@

Using data from a previous study of patients with amyotrophic lateral sclerosis (ALS)~\cite{sarica2017corticospinal}, we tested the performance of SGL in a
classification setting. The previous study predicted ALS status with a mean
accuracy of 80\% using a random forest algorithm based on \emph{a priori}
selection of features only within the CST bundle-of-interest.
SGL delivers improved predictive performance, with a cross-validated accuracy of {\alsAccuracy}\% and an area under the receiver operating characteristic curve (ROC AUC) of {\alsRocAuc}, without the need for \emph{a priori} feature engineering (Figure \cref{fig:class-results:probs}). In addition to this
classification performance, SGL also identifies the white matter tracts most
important for ALS classification. The relative importance of white matter
features is captured in the $\beta$ coefficients from \cref{supp:eq:sgl} of
the supplemental material \cite{supplement}.
\Cref{fig:class-results:tract-profiles} depicts these coefficients along the
right CST, plotted over the FA values for the control and ALS subject groups.
We find that SGL selects FA metrics in the corticospinal tract and
particularly in the right corticospinal tract as most important to ALS
classification, confirming previous findings \cite{van2011upper,
toosy2003diffusion, sarica2014tractography, sage2007quantitative,
sage2009quantitative, karlsborg2004corticospinal, ellis1999diffusion,
cosottini2005diffusion, ciccarelli2009investigation, abe2010voxel} and
identifying the portions of the brain that were selected \emph{a priori} in
the previous study from which we obtained the data
\cite{sarica2017corticospinal}.

The $\beta$ coefficients exhibit high bundle level sparsity; only some
bundles are important, which can be be confirmed by observing the value of
$\alpha$, the regularization hyperparameter that controls the mixture of the
group lasso and lasso penalties, selected through nested cross-validation
(see \cref{supp:eq:sgl} of the supplemental material \cite{supplement}).
If $\alpha$ is closer to zero, it indicates that the phenotype in question
preferentially correlates with only a few groups of covariates. For the ALS
dataset, $\alpha = $ \alsLRatio, confirming that the white matter correlates
of ALS reside mostly in one bundle, namely the CST.

Analyzing the ways in which the model mislabels individuals also provides
insight. We found that mislabelled subjects are outliers relative to their
group with respect to diffusion features of the CST (Figure
\Cref{fig:class-results:false-positives,fig:class-results:false-negatives}).
The false positive classifications have reduced FA in one or more of the three
sections of the CST where $\norm*{\hat{\beta}}$ is large
\cref{fig:class-results:tract-profiles}. The false negative subjects have FA
profiles that oscillate between the two group means. Thus, when the SGL
method predicts incorrectly this is done in a comprehensible manner.

@@

To test the performance of SGL with tractometry data in a continuous
regression task, we focus here on the prediction of biological age in three
datasets named WH, HBN, and Cam-CAN (see supplemental material
\cite{supplement} for a description of each dataset). Prediction of ``brain
age'' is a commonly undertaken task in neuroimaging machine learning, in part
because these predictions, and deviations therefrom, may be diagnostic of
overall brain health (for a review, see Cole et al.~\cite{Cole2019-rz}).
However, as Nelson et al.~\cite{nelson2019biomarkers} have observed, aging
biomarkers are subject to unique challenges and tend to be noncausative. Our
interest in aging here relies on its utility as a methodological benchmark.
Biological age operates on a natural scale, with meaningful and easily
understood units, and it's popularity as a machine learning target makes it
valuable for comparisons with other studies.

The WH \cite{yeatman2014lifespan}, HBN \cite{alexander2017open}, and Cam-CAN
\cite{shafto2014cambridge,taylor2017cambridge} datasets used here contain
data from 76, 978, and 640 subjects, respectively, ranging from 6-50, 5-21,
and 18-88 years of age, respectively. In each case, biological age was used
as the predicted variable $y$. SGL was fit to the tractometry-extracted
features FA and MD in 18 major brain tracts, with diffusion metrics extracted
from diffusion tensor imaging (DTI) for the WH dataset and diffusion kurtosis
imaging (DKI) \cite{jensen2005diffusion} for the HBN and Cam-CAN datasets.

To evaluate the fit of the model, we used a nested cross-validation
procedure. In this procedure, batches of subjects are held out. For each
batch (or fold), the model is fully fit without this data. Then, once the
parameters are fixed, the model is inverted to predict the ages of held out
subjects based on the linear coeffiecients. This scheme automatically finds
the right level of sparsity and fits the coefficients to the ill-posed linear
model, while guarding against overfitting. SGL accurately predicts the age of
the subjects in this procedure, with a median absolute error of {\whMae},
{\hbnMae}, and {\camcanMae} years for the WH, HBN, and Cam-CAN datasets,
respectively and coefficients of determination $R^2 = $ {\whRsq} , {\hbnRsq},
and {\camcanRsq}, respectively (see \cref{fig:age-results}, top panels). The
predictions for Cam-CAN are competitive with a recent state of the art
prediction \cite{mcpherson2020single}, which used streamline density to
estimate the brain's structural connectivity and achieved $R^2 = 0.63$. The
median absolute errors are also lower than the results of a recent study that
predicted age in a large sample that included the Cam-CAN data, and was based
on diffusion MRI features \cite{Richard2018-ux}.
